Commutative Character Sheaves

### 2015-11-15 followup.tex
Playing with subsubsections in Section 3. What do you think?

In order to get character sheaves that correspond to characters in $G(k)^*$, we must discard those
character sheaves whose restriction to the derived subgroup is nontrivial.
The following proposition will be proven in Section \ref{ssec:proofs}.
\begin{proposition} \label{prop:Gder_triv}
Suppose $G$ is a smooth group scheme and $\cs{L} \in \CS(G)$ is a linear character sheaf on $G$.
Then the restriction of $\cs{L}$ to $G\der$ is trivial if and only if $\cs{L} \cong q^*(\cs{L}\ab)$ for some
character sheaf $\cs{L}\ab$ on $G\ab$.
\end{proposition}

@@

\begin{theorem}\label{thm:Gab}
Pull-back along the abelianization $q : G \to G\ab$
defines an equivalence of categories
\[
\CCS(G\ab) \to \CCS(G).
\]
\end{theorem}
%
Sections~\ref{ssec:equivariant1}, \ref{ssec:equivariant2}, \ref{ssec:descent} and \ref{ssec:closed} set up machinery needed to prove Theorem~\ref{thm:Gab}.
The precise definition of the functor $\CCS(G\ab) \to \CCS(G)$ will be given with the proof of Theorem~\ref{thm:Gab}, in Section~\ref{ssec:proofs}.

@@

\subsection{Equivariant Weil local systems}\label{ssec:equivariant1}

@@

\subsection{Equivariant linear character sheaves}\label{ssec:equivariant2}

@@

\subsection{Descent along a torsor}\label{ssec:descent}

@@

\subsection{Quotient by a closed subgroup}\label{ssec:closed}

@@

\subsection{Proofs for results from Section \ref{ssec:noncomdef}}\label{ssec:proofs}

@@

\begin{proof}[Proof of Proposition~\ref{prop:Gder_triv}]
The abelianization $G\ab\ceq G/G\der$ exists in schemes over $\Fq$ \cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}.
Set $H=G\der$.
Let $j : H\hookrightarrow G$ be the inclusion and let $q : G \to G\ab$ be the quotient. 
With reference to \eqref{qH} and Section~\ref{ssec:closed}, consider the following diagram.
%\[
%\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
%1 \arrow{r} & H \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
%\end{tikzcd}
%\]
%gives
\[
\begin{tikzcd}
\CS(H) & \arrow{l}[swap]{j^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
 \CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} &  \CS_{H}(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]
%
Here we use the notation $\CS'(G)$ from Section~\ref{ssec:noncomdef} and the observation that $\CS'(G)$ is precisely the category $\CS^{H}(G)$ from Section~\ref{ssec:closed} for $H=G\der$.
The forget
Since the quotient $q : G \to G\ab$ is a trivial $H$-torsor in the fppf topology \cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}, 
it now follows from Lemma~\ref{lem:torsor} that the comparison functor $\CS(G\ab)\to \CS_{H}(G)$ is an equivalence.
On the other hand, by Lemma~\ref{lem:beta}, $\CS'(G) \to \CS_{H}(G)$, defined by $(\cs{L},\mu,\beta)\mapsto (\cs{L},\mu, \nu)$, is an equivalence.
Since the essential image of the forgetful functor $\CS'(G)\to \CS(G)$ is precisely the category of $(\cs{L},\mu)\in \CS(G)$ such that $\cs{L}\vert_{H} \iso (\EE)_{H}$, this completes the proof of Proposition~\ref{prop:Gder_triv}.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:Gab}] 
Here we use notation from the proof of Proposition~\ref{prop:Gder_triv};
in particular, $H = G\der$.
By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{H}(G)$ is an equivalence, it determines a full subcategory $\CCS_{H}(G)$ which is equivalent to $\CCS(G\ab)$, as pictured below.
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of the equivalence $\CS_{H}(G) \to \CS'(G)$ is $\CCS(G)$.
\[
\begin{tikzcd}[row sep=30, column sep=40]
\CS(H) 
& \arrow{l}[swap]{j^*} \CS(G) 
&\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
\CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} 
&  \CS_{H}(G) \arrow{u}{\text{forget}}  
&  %\arrow[bend left=35pt, dashed]{dll} 
\CCS(G\ab) \arrow{u}[swap]{\text{full sub}} \arrow{dl}{\text{equiv}} \\
\CCS(G) \arrow{u}{\text{full sub}} \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}}
& \arrow{u}{\text{full sub}}  \CCS_{H}(G) 
& 
\end{tikzcd}
\]

Suppose $(\cs{L},\mu,\beta)\in \CS'(G)$ and $(\cs{L},\mu,\nu) = q^*(\cs{L}\ab,\mu\ab)$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{sec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{H} \to (\EE)_{H}$ as in Section~\ref{ssec:noncomdef}.
Then the diagram in Definition~\ref{def:CCS} is precisely the result of applying the functor $(q\times q)^*$ to the diagram in Definition~\ref{def:CCScom}, as pictured below; 
in particular $\gamma = (q\times q)^* \xi$.
  \[
  \begin{tikzcd}[row sep=10, column sep = 20]
   m_0^*\cs{L}_0 \arrow{dd}[swap]{\xi} \arrow{r}{\mu_0} 
   & \cs{L}_0\boxtimes\cs{L}_0 \arrow{dd}{\vartheta}
&&&   m^*\cs{L} \arrow{dd}[swap]{\gamma} \arrow{r}{\mu} 
 & \cs{L}\boxtimes\cs{L} \arrow{dd}{\vartheta}\\
 && \arrow{r}{(q\times q)^*}  &\ & & &\\ 
   \theta^*(m_0^*\cs{L}_0) \arrow{r}{{\theta}^*\mu_0} 
 &  \theta^*(\cs{L}_0\boxtimes\cs{L}_0)
&&&   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} 
&  \theta^*(\cs{L}\boxtimes\cs{L})
  \end{tikzcd}
\]
By Lemma~\ref{lem:HH}, we may interpret the diagram on the right, above, as a diagram in $\Loc_{H\times H}(G\times G)$.
By Lemma~\ref{lem:torsor}, this corresponds to a diagram in $\Loc(G\ab\times G\ab)$, necessarily the diagram on the left, above, and also that the diagram in Definition~\ref{def:CCS} commutes if and only if the diagram in Definition~\ref{def:CCScom} commutes. 
In other words, $(\cs{L},\mu,\beta)\in \CCS(G)$
if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G)$.
This completes the proof of Theorem~\ref{thm:Gab}.
%This also defines the functor $\CCS(G\ab) \to \CCS(G)$ by the dashed arrow in the diagram above.
\end{proof}

>>>>

In order to get character sheaves that correspond to characters in $G(k)^*$, we must discard those
character sheaves whose restriction to the derived subgroup is nontrivial.
The following proposition will be proven in Section \ref{ssec:proof1}.
\begin{proposition} \label{prop:Gder_triv}
Suppose $G$ is a smooth group scheme and $\cs{L} \in \CS(G)$ is a linear character sheaf on $G$.
Then the restriction of $\cs{L}$ to $G\der$ is trivial if and only if $\cs{L} \cong q^*(\cs{L}\ab)$ for some
character sheaf $\cs{L}\ab$ on $G\ab$.
\end{proposition}

@@

\begin{theorem}\label{thm:Gab}
Pull-back along the abelianization $q : G \to G\ab$
defines an equivalence of categories
\[
\CCS(G\ab) \to \CCS(G).
\]
\end{theorem}
%
Sections~\ref{ssec:equivariant1}, \ref{ssec:equivariant2}, \ref{ssec:descent} and \ref{ssec:closed} set up machinery needed to prove Theorem~\ref{thm:Gab}.
The precise definition of the functor $\CCS(G\ab) \to \CCS(G)$ will be given with the proof of Theorem~\ref{thm:Gab}, in Section~\ref{ssec:proof2}.

@@

\subsection{Proofs for results from Section \ref{ssec:noncomdef}} \label{ssec:proofs}

\subsubsection{Equivariant Weil local systems}\label{ssec:equivariant1}

@@

\subsubsection{Equivariant linear character sheaves}\label{ssec:equivariant2}

@@

\subsubsection{Descent along a torsor}\label{ssec:descent}

@@

\subsubsection{Quotient by a closed subgroup}\label{ssec:closed}

@@

@@

\subsubsection{Proof of Proposition~\ref{prop:Gder_triv}}\label{ssec:proof1}

%\begin{proof}[Proof of Proposition~\ref{prop:Gder_triv}] 
The abelianization $G\ab\ceq G/G\der$ exists in schemes over $\Fq$ \cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}.
Set $H=G\der$.
Let $j : H\hookrightarrow G$ be the inclusion and let $q : G \to G\ab$ be the quotient. 
With reference to \eqref{qH} and Section~\ref{ssec:closed}, consider the following diagram.
%\[
%\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
%1 \arrow{r} & H \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
%\end{tikzcd}
%\]
%gives
\[
\begin{tikzcd}
\CS(H) & \arrow{l}[swap]{j^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
 \CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} &  \CS_{H}(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]
%
Here we use the notation $\CS'(G)$ from Section~\ref{ssec:noncomdef} and the observation that $\CS'(G)$ is precisely the category $\CS^{H}(G)$ from Section~\ref{ssec:closed} for $H=G\der$.
The forget
Since the quotient $q : G \to G\ab$ is a trivial $H$-torsor in the fppf topology \cite{SGA3}*{Expos\'e {VI}$_A$, Thm. 3.2}, 
it now follows from Lemma~\ref{lem:torsor} that the comparison functor $\CS(G\ab)\to \CS_{H}(G)$ is an equivalence.
On the other hand, by Lemma~\ref{lem:beta}, $\CS'(G) \to \CS_{H}(G)$, defined by $(\cs{L},\mu,\beta)\mapsto (\cs{L},\mu, \nu)$, is an equivalence.
Since the essential image of the forgetful functor $\CS'(G)\to \CS(G)$ is precisely the category of $(\cs{L},\mu)\in \CS(G)$ such that $\cs{L}\vert_{H} \iso (\EE)_{H}$, this completes the proof of Proposition~\ref{prop:Gder_triv}.
%\end{proof}

\subsubsection{Proof of Theorem~\ref{thm:Gab}}\label{ssec:proof2}

%\begin{proof}[Proof of Theorem~\ref{thm:Gab}] 
Here we use notation from the proof of Proposition~\ref{prop:Gder_triv};
in particular, $H = G\der$.
By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{H}(G)$ is an equivalence, it determines a full subcategory $\CCS_{H}(G)$ which is equivalent to $\CCS(G\ab)$, as pictured below.
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of the equivalence $\CS_{H}(G) \to \CS'(G)$ is $\CCS(G)$.
\[
\begin{tikzcd}[row sep=30, column sep=40]
\CS(H) 
& \arrow{l}[swap]{j^*} \CS(G) 
&\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
\CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} 
&  \CS_{H}(G) \arrow{u}{\text{forget}}  
&  %\arrow[bend left=35pt, dashed]{dll} 
\CCS(G\ab) \arrow{u}[swap]{\text{full sub}} \arrow{dl}{\text{equiv}} \\
\CCS(G) \arrow{u}{\text{full sub}} \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}}
& \arrow{u}{\text{full sub}}  \CCS_{H}(G) 
& 
\end{tikzcd}
\]

Suppose $(\cs{L},\mu,\beta)\in \CS'(G)$ and $(\cs{L},\mu,\nu) = q^*(\cs{L}\ab,\mu\ab)$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{sec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{H} \to (\EE)_{H}$ as in Section~\ref{ssec:noncomdef}.
Then the diagram in Definition~\ref{def:CCS} is precisely the result of applying the functor $(q\times q)^*$ to the diagram in Definition~\ref{def:CCScom}, as pictured below; 
in particular $\gamma = (q\times q)^* \xi$.
  \[
  \begin{tikzcd}[row sep=10, column sep = 20]
   m_0^*\cs{L}_0 \arrow{dd}[swap]{\xi} \arrow{r}{\mu_0} 
   & \cs{L}_0\boxtimes\cs{L}_0 \arrow{dd}{\vartheta}
&&&   m^*\cs{L} \arrow{dd}[swap]{\gamma} \arrow{r}{\mu} 
 & \cs{L}\boxtimes\cs{L} \arrow{dd}{\vartheta}\\
 && \arrow{r}{(q\times q)^*}  &\ & & &\\ 
   \theta^*(m_0^*\cs{L}_0) \arrow{r}{{\theta}^*\mu_0} 
 &  \theta^*(\cs{L}_0\boxtimes\cs{L}_0)
&&&   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} 
&  \theta^*(\cs{L}\boxtimes\cs{L})
  \end{tikzcd}
\]
By Lemma~\ref{lem:HH}, we may interpret the diagram on the right, above, as a diagram in $\Loc_{H\times H}(G\times G)$.
By Lemma~\ref{lem:torsor}, this corresponds to a diagram in $\Loc(G\ab\times G\ab)$, necessarily the diagram on the left, above, and also that the diagram in Definition~\ref{def:CCS} commutes if and only if the diagram in Definition~\ref{def:CCScom} commutes. 
In other words, $(\cs{L},\mu,\beta)\in \CCS(G)$
if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G)$.
This completes the proof of Theorem~\ref{thm:Gab}.
%This also defines the functor $\CCS(G\ab) \to \CCS(G)$ by the dashed arrow in the diagram above.
%\end{proof}
